This thesis tackles the problem of testing (legacy) production
systems such as those of our industrial partner Michelin, one of
the three largest tire manufacturers in the world, by means of
Model-based Testing. A production system is defined as a set of
production machines controlled by a software, in a factory.
Despite the large body of work within the field of Model-based
Testing, a common issue remains the writing of models describing
either the system under test or its specification. It is a
tedious task that should be performed regularly in order to keep
the models up to date (which is often also true for any
documentation in the Industry). A second point to take into
account is that production systems often run continuously and
should not be disrupted, which limits the use of most of the
existing classical testing techniques.

We present an approach to infer exact models from traces,
\emph{i.e.} sequences of events observed in a production
environment, to address the first issue. We leverage the data
exchanged among the devices and software in a black-box
perspective to construct behavioral models using different
techniques such as expert systems, model inference, and machine
learning. It results in large, yet partial, models gathering the
behaviors recorded from a system under analysis. We introduce a
context-specific algorithm to reduce such models in
order to make them more usable while preserving trace equivalence
between the original inferred models and the reduced ones. These
models can serve different purposes, \emph{e.g.}, generating
documentation, data mining, but also testing.

To address the problem of testing production systems without
disturbing them, this thesis introduces an \emph{offline passive
Model-based Testing technique}, allowing to detect differences
between two production systems. This technique leverages the
inferred models, and relies on two implementation relations: a
slightly modified version of the existing trace preorder
relation, and a weaker implementation proposed to overcome the
partialness of the inferred models.

Overall, the thesis presents \emph{Autofunk}, a modular framework
for model inference and testing of production systems, gathering
the previous notions. Its Java implementation has been applied to
different applications and production systems at Michelin, and
this thesis gives results from different case studies. The
prototype developed during this thesis should become a standard
tool at Michelin.

\textbf{Keywords.} Model inference, expert system, production
system, offline passive testing, conformance.
