% !TEX root = ../../thesis.tex
%
\chapter{Testing applied to production systems}
\label{sec:testing}

In this chapter, we tackle the problem of testing production
systems (as defined in the previous chapter), without disturbing
them, and without having any specification \emph{a priori}. Those
two constraints sounds familiar as they have already been taken
into consideration in the previous chapter. We present a passive
testing technique built on-top of \emph{Autofunk v3}. First, we
infer reference models of a system under analysis $\mathit{Sua}$
using the technique presented in the previous chapter.
Then, we check whether a system under test $\mathit{Sut}$
conforms to these inferred models by means of two implementation
relations. We use a slightly modified version of the trace
preorder relation to strictly check conformance between the two
systems. Because the inferred models are partial (as explained in
the previous chapter), they likely lack information, which
implies that a strict relation may be too strong. That is why we
propose our own (weaker) implementation relation to comply with
Michelin's needs.  Such a relation is less strict than the first
one in order to accept non-standard behaviors down to a certain
point. These two relations are leveraged in a scalable algorithm
for performing offline passive conformance testing.  We end this
chapter with a few results on this testing technique.\\

\minitoc

\pagebreak

\input{src/testing/1_introduction}
\input{src/testing/2_offline-passive-testing}
\input{src/testing/3_experimentation}
\input{src/testing/4_conclusion}

\cleardoublepage
\blankpage
