\subsection{Offline passive testing algorithm}
\label{sec:testing:offline}

Our offline passive testing algorithm, which aims to check
whether the two previous implementation relations hold, is given
in Algorithm \ref{algo:check}. It takes the complete traces of
$\mathit{Sut}$, as well as the models $R(\EuScript{S}^{N})$ and
$D(\EuScript{S}^{N})$, with regards to Proposition
\ref{rel:impl12} and Proposition \ref{rel:impl2}. It returns the
verdict "Pass$\leq_{ct}$" if the relation $\leq_{ct}$ is
satisfied, and the verdict "Pass$\leq_{mct}$" if $\leq_{mct}$ is
satisfied.

Algorithm \ref{algo:check} relies upon the function ${check(Trace
~trace, STS ~\EuScript{S})}$ to check whether the trace $trace =
(a_1(p), \alpha_1) \dots (a_m(p), \alpha_m)$ is a trace of the STS
$\EuScript{S}$. If a STS path $b$ is composed of the same
sequence of symbols than $trace$ (line
\ref{algo:check:line:exists}), the function tries to find a
matrix column (\emph{i.e.} a vector) $M = M_{[b]}[*,c_{[b]}]$ ($1
\leq c_{[b]} \leq k$) such that every variable assignment
$\alpha_j$ satisfies the guard $M[j]$. If such a column of guards
exists, the function returns $True$, and $False$ otherwise.

First, this algorithm covers every trace $trace$ of
$CTraces({Sut})$, and tries to find a STS $R(\EuScript{S}_i^{N})$
such that $trace$ is also a trace of $R(\EuScript{S}_i^{N})$ with
$check(trace, R(\EuScript{S}_i^{N}))$ (line
\ref{algo:check:line:check1}).  If no model
$R(\EuScript{S}_i^{N})$ is found, $trace$ is added to the set
$T_1$ (line \ref{algo:check:line:t1}), which gathers the possibly
fail traces with respect to $\leq_{ct}$.

Thereafter, this algorithm performs the same step but using the
STS $D(\EuScript{S}^{N})$ (line \ref{algo:check:line:check2}).
One more time, if no model $D(\EuScript{S}_i^{N})$ is found, the
trace $trace$ is added to the set $T_2$ (line
\ref{algo:check:line:t2}), which gathers the possibly fail
traces with respect to the relation $\leq_{mct}$.

Finally, if $T_1$ is empty, the verdict "Pass$\leq_{ct}$" is
returned, which means that the first implementation relation
holds. Otherwise, $T_1$ is provided. If $T_2$ is empty, the
verdict "Pass$\leq_{mct}$" is returned. Otherwise $T_2$ is
returned.

\begin{algorithm}[h]
    \SetKwInOut{Input}{Input}
    \SetKwInOut{Output}{Output}
    \SetKwFunction{check}{check}

    \Input{
        $R(\EuScript{S}^{N}),
        D(\EuScript{S}^{N}),
        CTraces({Sut})$
    }
    \Output{Verdicts and/or possibly fail trace sets $T_1, T_2$ }

    BEGIN\;

    $T_1 = \emptyset$\;
    $T_2 = \emptyset$\;

    \ForEach{$trace \in CTraces({Sut})$}{

        \For{$i = 1, \dots, n$}{
            \If{\check($trace$,
            $R(\EuScript{S}_i^{N})$)}{\label{algo:check:line:check1}
                break\;
            }
        }

        \If{$i == n$} {
            $T_1=T_1 \cup \{trace\}$\;\label{algo:check:line:t1}

            \For{$i = 1, \dots, n$}{

                \If{\check($trace$,
                $D(\EuScript{S}_i^{N})$)}{\label{algo:check:line:check2}
                    break\;
                }
            }

            \If{$i==n$} {
                $T_2=T_2 \cup \{trace\}$\;\label{algo:check:line:t2}
            }
        }
    }%endfor1

    \BlankLine

    \If{$T_1==\emptyset$}{
        return "Pass$\leq_{ct}$"\;}
    \Else{
        \If{$T_2==\emptyset$}{
            return "Pass$\leq_{mct}$" and $T_1$\;
        }\Else{
            return $T_1$ and $T_2$\;
        }
    }

    END\;

    \BlankLine
    \BlankLine

    \SetKwProg{Fn}{Function}{ is}{end}
    \Fn{check(Trace $trace$, STS $\EuScript{S}$) : boolean}{
        \If{$\exists ~b=l0_{\EuScript{S}}
            \xRightarrow{(a_1(p_1),G_{1},A_{1}) \dots
            (a_n(p_n),G_{n},A_{n})} l_n \mid trace =
            (a_1(p), \alpha_1)\dots (a_n(p), \alpha_n)$ and $l_n \in
            Pass$}{\label{algo:check:line:exists}

                $M_{[b]} = Mat(b)$ is the matrix $n \times k$ of $b$\;
                $c_{[b]} = 1$\;
                \While{$c_{[b]} \leq k$}{
                    $M = M_{[b]}[*,c_{[b]}]$\;

                    \For{$j = 1, \dots, n$}{
                        \If{$\alpha_j \not\models M[j]$}{
                            break\;
                        }
                    }

                    \If{$j == n$}{
                        return $True$\;
                    }
                    $c_{[b]}++$\;
                }
        }
        return $False$\;
    }

    \caption{Offline passive testing algorithm}
    \label{algo:check}
\end{algorithm}

When one of the implementation relations does not hold, this
algorithm offers the advantage of providing the possibly fail
traces of $CTraces({Sut})$. Such traces can be later analyzed to
check if $\mathit{Sut}$ is correct or not. That is very helpful
for Michelin engineers because it allows them to only focus on
what are potentially faulty behaviors, reducing debugging time,
and making engineers more efficient.

In the same spirit, our online passive testing algorithm is given
in the next section.

\clearpage
